---
title: "Genetic drift"
date: "October, 2018"
author:
  - name: M Perez-Enciso
    email: miguel.perez@uab.es
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '1'
    toc_float: true
  pdf_document:
    toc: yes
    toc_depth: 1
  theme: hpstr
  md_document:
    variant: markdown_github
  highlight: github---
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
    toc: yes
    toc_depth: '1'
  word_document:
    toc: yes
    toc_depth: '1'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Some popular R packages for population genetics
Population genetics in R

http://grunwaldlab.github.io/Population_Genetics_in_R/

PopGenome: An Efficient Swiss Army Knife for Population Genomic Analyses

https://cran.r-project.org/web/packages/PopGenome/

# Simple genetic drift program
### Before starting, think what is genetic drift?
### 
```{r, include=TRUE,warning=FALSE,echo=TRUE,eval=TRUE}
#----- simulates drift
# N is population size and p is initial allele frequency
drift <- function(N,p) {
   g=as.numeric()
   t=0
   while(p>0 & p<1) {
      genotypes <- rbinom(N,1,p)
      p<-mean(genotypes)
      t=t+1
      # the last column of each row contains current allele frequency
      genotypes=cbind(t(genotypes),p)
      g=rbind(g,genotypes)
   }
   return(g[,N+1])
}
```

EXERCISE: Guess what the program does and why

HINT:  look for help with 'help(command)' or '? command'

HELP in R: https://www.r-project.org/help.html

## Running the drift R function

```{r  include=TRUE,warning=FALSE,echo=TRUE,eval=TRUE}
# num of individuals
N=20
# initial allele frequency
p0=.2
f=drift(N,p0)
```

```{r  include=TRUE,warning=FALSE,echo=TRUE,eval=TRUE}
# list of colors for plotting each replicate
color = grDevices::colors()[grep('gr(a|e)y', grDevices::colors(), invert = T)]
plot(f, type='l', ylim=c(0,1), xlim=c(1,30), xlab='generation', ylab='f')
for (rep in seq(100)) {
    f = drift(N, p0)
    lines(f, col=sample(color,1))
}
```

EXERCISE: try different sizes (N) an dinitial frequencies (p)

EXERCISE: What is the probability of fixation of allele 1 ?

